---
title: "HeadSetSportsSummaryGenerator"
output: html_document
---

# README: 
# HOW TO USE THIS PROGRAM

#################################################
# STARTUP
## SIMPLIFIED: 
- Go to the drop down menu next to the run button at the top right of the code window
- Click on "Restart R and Run All Chunks"
- Continue to "DURING THE PROGRAM"

## STEPS 
### Step 1: Run the Libraries chunk
- Press the green arrow at the top right of the chunk called Libraries

### Step 2: Run the Functions chunk
- Press the green arrow at the top right of the chunk called Functions

### Step 3: Run the Main chunk
- Press the green arrow at the top right of the chunk called Main
- Continue to "DURING THE PROGRAM"
#################################################

# DURING THE PROGRAM
- A file selection box will pop up. Navigate to and select the file you want to clean. 
- The program will performing all the data cleaning
- A box will pop up asking for you to input the team name
  - Enter the team name without spaces
  - Whatever name you enter will be the start of the file name, for example
  - Entering "SpruceGrove" results in a final file name of "SpruceGroveCleaned.xls"
- A final box will pop up asking if you have another file to clean
  - Select an option by clicking on the option you want and clicking "Ok"
    - If you select "Yes", the program will return to the start of the "DURING THE PROGRAM"
    - If you select "No", the program will end

```{r Libraries, message = FALSE}
library(here)
library(tidyverse)
library(xlsx)
library(readxl)
library(lubridate)
library(rstudioapi)
library(svDialogs)
```


```{r Functions}
# Generate the overall data summary (first sheet of the document). 
# Returns the summary dataset
GenerateOverallSummary <- function(FileLocation) {
  # Read in the first sheet of the file which has the overall summaries of the player usage
  data <- read.xlsx(FileLocation, header = TRUE, sheetIndex = 1, startRow = 7) %>%
    # Split the task count box into several columns by activity
    separate(col = 2, into = c("Assessments", "Weekly_Updates", "WOWs", "Events", "Pre_Competition_Events", "Debriefs", "Videos"), sep = '\\|') %>% 
    # Extract the numeric value for activities completed by removing the text and saving the activity type as the column header
    mutate(Assessments = as.numeric(gsub("Assessments:", "", Assessments))) %>%
    mutate(Weekly_Updates = as.numeric(gsub("Weekly Updates:", "", Weekly_Updates))) %>%
    mutate(WOWs = as.numeric(gsub("WOWs:", "", WOWs))) %>%
    mutate(Events = as.numeric(gsub("Events:", "", Events))) %>%
    mutate(Pre_Competition_Events = as.numeric(gsub("Pre Competition Tasks: ", "", Pre_Competition_Events))) %>%
    mutate(Debriefs = as.numeric(gsub("Debriefs:", "", Debriefs))) %>%
    mutate(Videos = as.numeric(gsub("Videos: ", "", Videos)))
  
  # Return the overall performance summary
  return(data)
}


# Generate the long data set.
# Returns the long data set
GenerateLongDataSet <- function(FileLocation){
  # Read in each of the personal performance sheets and combine into one long list
  # The sheet names are also the player names
  SheetNames <- excel_sheets(FileLocation)
  
  # A counter just to keep track of how many sheets actually contain data.
  # It is used to identify the first usable sheet encountered
  GoodSheets <- 0
  
  # Read each of the sheets and combine them into a long list
  # We skip the first sheet since that contains the summary data
  for(sheet in 2:length(SheetNames)) {
    # Read in the whole sheet to check how many rows it has
    RowCheck <- read.xlsx(FileLocation, sheetIndex = sheet, headers = TRUE)
    
    # Check if the sheet has anything on it. If not, skip it
    if(length(RowCheck) == 0){
      next
    }
    
    # If the sheet has less than 6 rows it means the player didn't do any activities. Skip that sheet.
    if(dim(RowCheck)[1] < 6){
      next
    }
    
    # Check if this is the first sheet with usable data we have encountered
    # If so, generate the long dataset. 
    if(GoodSheets == 0){
    #  Read the sheet and input the column headers along with the player data
      DataLong <- read.xlsx(FileLocation, sheetIndex = sheet, startRow = 9) %>%
        head(-1) %>%
        add_column(Player = SheetNames[sheet], .before = "Date")
      GoodSheets <- GoodSheets + 1
      next
    }
    
    # For all other sheets of players that did something, input the data to a temporary dataset
    SheetData <- read.xlsx(FileLocation, sheetIndex = sheet, startRow = 9) %>%
      head(-1) %>%
      add_column(Player = SheetNames[sheet], .before = "Date")
    
    # Append the new data to the end of the full data set
    DataLong <- DataLong %>%
      rbind(SheetData)
    
    # Increment the number of sheets that had usable data
    GoodSheets <- GoodSheets + 1
  }
  
  # Convert the character dates to date data types
  DataLong$Date <- as.Date(DataLong$Date, "%m/%d/%Y")
  
  # Return the long data set
  return(DataLong)
}


# Calculates the time someone used HeadSet. 
# Returns a list with [[Months], [Days]].
# Time used is determined as time between the first and last assessment
TimeUsed <- function(Name) {
  # Get all the rows for the player
  PlayerInfo <- DataLong[which(DataLong$Player == Name), ]
  Dates <- PlayerInfo[which(PlayerInfo$Type == "assessment"), 2]
  # If they don't have 2 assessments, I label the time used as 0
  if(length(Dates) < 2) {
    return(list(0, 0))
  }
  # CALCULATE DAYS EACH PERSON USED THE PROGRAM
  # Start date is the first assessment date
  StartDate <- ISOdate(year(Dates[1]), month(Dates[1]), day(Dates[1]))
  # End date is the last assessment date
  EndDate <- ISOdate(year(Dates[length(Dates)]), month(Dates[length(Dates)]), day(Dates[length(Dates)]))
  # Find the days used
  DaysUsed <- as.numeric(EndDate - StartDate)
  # Find the months used
  MonthsUsed <- round(DaysUsed/30)
  # Combine days and months into a list
  TimeUsed <- list(DaysUsed, MonthsUsed)
  # Return the list of days used and months used
  return(TimeUsed)
}  


# Generate the dataset of total activities completed by each participant
# Returns the TotalActivities dataset
GenerateTotalActivities <- function(SummaryDataSet){
  # Pick the columns we want to include
  Summary <- SummaryDataSet %>% 
    select(c("Member.Name":"Videos"))
  
  # Rename the member column to "Player"
  colnames(Summary)[1] <- "Player"
  
  # Declare the vectors used to store time usage data
  MonthsUsed <- integer(length(Summary$Player))
  DaysUsed <- integer(length(Summary$Player))
  
  # Assign the time used for each player in respective vectors
  for (n in 1:length(Summary$Player)) {
    Times <- TimeUsed(Summary$Player[n])
    MonthsUsed[n] <- Times[[2]]
    DaysUsed[n] <- Times[[1]]
  }
  
  # Add the months and days used vectors to the descriptive dataset
  Summary <- Summary %>% 
    mutate(Months = MonthsUsed) %>% 
    mutate(Days = DaysUsed)
  
  # Return the descriptive summary data set
  return(Summary)
}


# Generate the dataset of activities completed before each assessment by each participant 
# Returns the ActivitiesPerAssessment dataset 
GenerateActivitiesPerAssessment <- function(LongDataSet, SummaryData) {
  
  ###############################################################################################
  # Make the new data set and establish the general structure
  
  # Generate new dataset
  Specifics <- LongDataSet[which(LongDataSet$Type == "assessment"), ] %>% 
    # Keep only the player name, the assessment date, and the scores
    select(Player, Date, Awareness:Motivation, Total.Points) %>%
    # Add the new columns for activities completed before each assessment
    mutate(Number_of_WOWs = 0, Number_of_Weeklys = 0, Number_of_Videos = 0, Number_of_Events = 0, Number_of_Debriefs = 0, Number_of_PreComp = 0) %>% 
    # Change the Total.Points column to be the assessment score and relocate it to be right after the date
    mutate(Assessment_Score = Total.Points, .after = "Date") %>%
    select(-Total.Points) %>%
    # Add a new column for the individual player assessment number
    add_column(Assessment_Number = 0, .after = "Player") 
  
  # Keep only the rows of players for which we have data according to the previously generated descriptive summary
  Specifics <- Specifics[which(Specifics$Player %in% SummaryData$Player), ]
  
  ################################################################################################
  # Find the tasks before each assessment
  
  # For each user (loop 1) we will go through each task (loop 2) and see how many of each event were completed before an assessment
  
  # Use this to keep track of the total assessments, and thus the row in the data set the counts should go in
  AssessmentCount <- 0
  for (n in 1:length(SummaryData$Player)) {
    # Get the player name
    PlayerName <- SummaryData$Player[n]
    # Get all tasks that player completed
    PlayerTasks <- LongDataSet[which(LongDataSet$Player == PlayerName), 3]
    # If they performed no tasks, skip them
    if(length(PlayerTasks) == 0) {
      next
    }
    # Set each task counter to 0
    WOWs = 0
    Weeklys = 0
    Videos = 0
    Events = 0
    Debriefs = 0
    PreComp = 0
    PlayerAssessmentCounter <- 0
    
    # Go through each task, incrementing the correct task counter
    # When it reaches an assessment it will put each task counter into the data set then reset the counters
    for (m in 1:length(PlayerTasks)) {
      if(PlayerTasks[m] == "wow") {
        WOWs = WOWs + 1
        }
      else if(PlayerTasks[m] == "weekly") {
        Weeklys = Weeklys + 1
        }
      else if(PlayerTasks[m] == "video") {
        Videos = Videos + 1
        }
      else if(PlayerTasks[m] == "event") {
        Events = Events + 1
        }
      else if(PlayerTasks[m] == "debrief") {
        Debriefs = Debriefs + 1
        }
      else if(PlayerTasks[m] == "Pre Competition Task") {
        PreComp = PreComp + 1
      }
      # When the task is an assessment, enter in the data set how many of each task was completed
      else if(PlayerTasks[m] == "assessment") {
        AssessmentCount = AssessmentCount + 1
        PlayerAssessmentCounter = PlayerAssessmentCounter + 1
        # Assign the total counts to the data set
        Specifics$Number_of_WOWs[AssessmentCount] <- WOWs
        Specifics$Number_of_Weeklys[AssessmentCount] <- Weeklys
        Specifics$Number_of_Videos[AssessmentCount] <- Videos
        Specifics$Number_of_Events[AssessmentCount] <- Events
        Specifics$Number_of_Debriefs[AssessmentCount] <- Debriefs
        Specifics$Number_of_PreComp[AssessmentCount] <- PreComp
        Specifics$Assessment_Number[AssessmentCount] <- PlayerAssessmentCounter
        # Reset each task counter
        WOWs = 0
        Weeklys = 0
        Videos = 0
        Events = 0
        Debriefs = 0
        PreComp = 0
      }
      # If there's a data error, it'll just skip it
      else {
        next
      }
    }
  }
  
  # Return the descriptive specifics data set
  return(Specifics)
}


# Generate the dataset of activities completed before each game debrief by each participant 
# Returns the ActivitiesPerGameDebrief dataset 
GenerateActivitiesPerGameDebrief <- function(LongDataSet, SummaryData) {
  
  ###############################################################################################
  # Make the new data set and establish the general structure
  
  # Generate new dataset
  Specifics <- LongDataSet[which(LongDataSet$Type == "debrief" & substr(LongDataSet$Description, 1, 4) == "Game"), ] %>% 
    # Keep only the player name, the assessment date, and the scores
    select(Player, Date, Total.Points, ) %>%
    # Add the new columns for activities completed before each assessment
    mutate(Number_of_Assessments = 0,Number_of_WOWs = 0, Number_of_Weeklys = 0, Number_of_Videos = 0, Number_of_Events = 0,  Number_of_PreComp = 0) %>% 
    # Change the Total.Points column to be the game debrief score and relocate it to be right after the date
    mutate(Game_Debrief_Score = Total.Points, .after = "Date") %>%
    select(-Total.Points) %>%
    # Add a new column for the individual player game debrief number
    add_column(Game_Debrief_Number = 0, .after = "Date") 
  
  # Keep only the rows of players for which we have data according to the previously generated descriptive summary
  Specifics <- Specifics[which(Specifics$Player %in% SummaryData$Player), ]
  
  ################################################################################################
  # Find the tasks before each assessment
  
  # For each user (loop 1) we will go through each task (loop 2) and see how many of each event were completed before an assessment
  
  # Use this to keep track of the total assessments, and thus the row in the data set the counts should go in
  TotalGameDebriefCount <- 0
  for (n in 1:length(SummaryData$Player)) {
    # Get the player name
    PlayerName <- SummaryData$Player[n]
    # Get all tasks that player completed
    PlayerTasks <- LongDataSet[which(LongDataSet$Player == PlayerName), 3]
    TaskDescriptions <- LongDataSet[which(LongDataSet$Player == PlayerName), 4]
    # If they performed no tasks, skip them
    if(length(PlayerTasks) == 0) {
      next
    }
    # Set each task counter to 0
    Assessments = 0
    WOWs = 0
    Weeklys = 0
    Videos = 0
    Events = 0
    PreComp = 0
    DebriefPractice = 0
    PlayerGameDebriefCounter <- 0
    
    # Go through each task, incrementing the correct task counter
    # When it reaches an assessment it will put each task counter into the data set then reset the counters
    for (m in 1:length(PlayerTasks)) {
      if(PlayerTasks[m] == "assessment") {
        Assessments = Assessments + 1
      }
      else if(PlayerTasks[m] == "wow") {
        WOWs = WOWs + 1
        }
      else if(PlayerTasks[m] == "weekly") {
        Weeklys = Weeklys + 1
        }
      else if(PlayerTasks[m] == "video") {
        Videos = Videos + 1
        }
      else if(PlayerTasks[m] == "event") {
        Events = Events + 1
        }
      else if(PlayerTasks[m] == "Pre Competition Task") {
        PreComp = PreComp + 1
      }
      else if((PlayerTasks[m] == "debrief") & (substr(TaskDescriptions[m], 1, 8) == "Practice")) {
        DebriefPractice = DebriefPractice + 1
      }
      # When the task is an assessment, enter in the data set how many of each task was completed
      else if((PlayerTasks[m] == "debrief") & (substr(TaskDescriptions[m], 1, 4) == "Game")) {
        TotalGameDebriefCount = TotalGameDebriefCount + 1
        PlayerGameDebriefCounter = PlayerGameDebriefCounter + 1
        # Assign the total counts to the data set
        Specifics$Number_of_Assessments[TotalGameDebriefCount] <- Assessments
        Specifics$Number_of_WOWs[TotalGameDebriefCount] <- WOWs
        Specifics$Number_of_Weeklys[TotalGameDebriefCount] <- Weeklys
        Specifics$Number_of_Videos[TotalGameDebriefCount] <- Videos
        Specifics$Number_of_Events[TotalGameDebriefCount] <- Events
        Specifics$Number_of_PreComp[TotalGameDebriefCount] <- PreComp
        Specifics$Number_of_Debrief_Practices[TotalGameDebriefCount] <- DebriefPractice
        Specifics$Game_Debrief_Number[TotalGameDebriefCount] <- PlayerGameDebriefCounter
        # Reset each task counter
        Assessments = 0
        WOWs = 0
        Weeklys = 0
        Videos = 0
        Events = 0
        PreComp = 0
        DebriefPractice = 0
      }
      # If there's a data error, it'll just skip it
      else {
        next
      }
    }
  }
  
  # Return the descriptive specifics data set
  return(Specifics)
}
```


```{r Main}
repeat{
  # Insert the file location here
  FileLocation <- selectFile(caption = "Select the HeadSet team data file", 
                     filter = "XLS Files (*.xls)", 
                     path = here())
  
  # I'm not sure if the selectFile function is compatible across platforms (it works on Windows)
  # If you have issues picking the file, comment out the selectFile() line and uncomment the one below
  # Insert the file path to the file you want to evaluate
  # FileLocation <- here("Spruce Grove Saints Junior Hockey_Overall Points_09_01_2020_05_29_2021.xls")
  
  # Generate the overall summary dataset
  DataSummary <- GenerateOverallSummary(FileLocation)
  
################################################################################
###################### GENERATE THE DATASETS FOR THE FILE ######################
################################################################################
  
  # Generate the long data set with all players that performed at least 1 activity. 
  # This is the complete list of activities performed by each player
  DataLong <- GenerateLongDataSet(FileLocation)
  
  # Generate the Total Activities dataset
  # This is the summary of each player (how many of each task they completed and the time they used the program)
  TotalActivities <- GenerateTotalActivities(DataSummary)
  
  # Generate the Activities per Assessment dataset
  # This is the data set that shows how many of each task was completed before each assessment, and the assessment scores
  ActivitiesPerAssessment <- GenerateActivitiesPerAssessment(DataLong, TotalActivities)
  
  # Generate the Activities per Game Debrief dataset
  # This is the data set that shows how many of each task was completed before each game debrief, and the debrief scores
  ActivitiesPerGameDebrief <- GenerateActivitiesPerGameDebrief(DataLong, TotalActivities)
  
################################################################################
  
  # This gets the team name from the user and will make the entered team name the file name
  teamName <- dlgInput("Enter the team name (no spaces). This will be part of the file name.", Sys.info()["user"])$res
  FileName <- paste(teamName, "Cleaned.xls", sep = "")
  
  
  
  # This writes the generated data sets to one excel file on different sheets
  write.xlsx(TotalActivities, FileName, sheetName = "Total Activities", row.names = FALSE, showNA = FALSE)
  write.xlsx(ActivitiesPerAssessment, FileName, sheetName = "Activities per Assessment", row.names = FALSE, append = TRUE, showNA = FALSE)
  write.xlsx(ActivitiesPerGameDebrief, FileName, sheetName = "Activities per Game Debrief", row.names = FALSE, append = TRUE, showNA = FALSE)
  write.xlsx(DataLong, FileName, sheetName = "Long Dataset", row.names = FALSE, append = TRUE, showNA = FALSE)

  
  
  # This checks whether or not the user wants to clean another dataset
  # If the response is yes, the main chunk will repeat
  # If the response is no (or anything else), then the code will end
  AnotherFile <- dlg_list(choices = c("No", "Yes"), title = "Another file?")$res
  if(AnotherFile == "Yes") {
    next
  }
  else {
    break
  }
}
```




