---
title: "OverallDataCondenser"
output: html_document
---

### READ ME:

# This program works best if all the file you wish to condense are in one directory. 
# The way the file selection works, you can only select files from one directory at a time, so if you want to combine data from multiple folders, 
#     you'll have to run this multiple times. It's easiest just to copy/move the individual data files into one folder so you can combine them all in one go.


```{r Libraries, message=FALSE}
library(tcltk)
library(tidyverse)
library(xlsx)
library(svDialogs)
```



```{r CombineData}
####### READ ME #########
# Choose all the files to condense
# If you are combining a "pre-condensed" list and other files, put the "pre-condensed" file first
#########################
file_list <- tk_choose.files()

# Condense the files
first_file = TRUE
for(file in file_list) {
  
  ######### ONLY APPLIES TO FIRST FILE EACH USE ##########
  # If it's the first file, generate the overall condensed data sets
  if (first_file) {
    # Ge the summary and specifics data
    summary <- read.xlsx(file, sheetIndex = 1)
    assessments <- read.xlsx(file, sheetIndex = 2)
    gameDebriefs <- read.xlsx(file, sheetIndex = 3)
    
    # If the first file is the condensed file set, just make that the base condensed data set
    if (colnames(summary)[1] == "Team_Name") {
      condensed_summary <-  summary
      condensed_assessments <- assessments
      condensed_game_debriefs <- gameDebriefs
      next
    }
    
    # If the first file is just a cleaned dataset (not the condensed one)
    # Extract the team name from the file name
    team_name <- sub("Cleaned.*", "", basename(file))
    # If the first file is an uncondensed dataset, make the condensed datasets
    condensed_summary <-  summary %>% 
      add_column(Team_Name = team_name, .before = "Player")
    # Make new activities per assessments dataset
    condensed_assessments <- assessments %>% 
      add_column(Team_Name = team_name, .before = "Player")
    # Make new activities per game debrief dataset
    condensed_game_debriefs <- gameDebriefs %>% 
      add_column(Team_Name = team_name, .before = "Player")
    # Mark that the first file has been used
    first_file <- FALSE
    next
  }
  #################################################
  
  
  ##### ALL OTHER ITERATIONS ###############
  
  # For all other files, make new data sets and combine them with the overall condensed set
  # Extract the team name from the file name
  team_name <- sub("Cleaned.*", "", basename(file))
  # Make new summary data set
  new_summary <- read.xlsx(file, sheetIndex = 1) %>%
    add_column(Team_Name = team_name, .before = "Player")
  # Make new activities per assessment dataset
  new_assessments <- read.xlsx(file, sheetIndex = 2) %>%
    add_column(Team_Name = team_name, .before = "Player")
  # Make new activities per game debrief dataset
  new_game_debriefs <- read.xlsx(file, sheetIndex = 3) %>%
    add_column(Team_Name = team_name, .before = "Player")
  
  # Combine the new datasets with the overall condensed set
  condensed_summary <- condensed_summary %>%
    rbind(new_summary)
  condensed_assessments <- condensed_assessments %>%
    rbind(new_assessments)
  condensed_game_debriefs <- condensed_game_debriefs %>%
    rbind(new_game_debriefs)
}


# This gets the file name from the user
FileName <- dlgInput("Enter the file name for the condensed data (ie. CondensedData): ", Sys.info()["user"])$res
FileName <- paste(FileName, ".xls", sep = "")
# Write the datasets to a new .xls file
write.xlsx(condensed_summary, FileName, sheetName = "Summary", row.names = FALSE, showNA = FALSE)
write.xlsx(condensed_assessments, FileName, sheetName = "Activities per Assessment", row.names = FALSE, append = TRUE, showNA = FALSE)
write.xlsx(condensed_game_debriefs, FileName, sheetName = "Activites per Game Debrief", row.names = FALSE, append = TRUE, showNA = FALSE)
```

