---
title: "HeadSet WebScraper"
author: "RJ Cass"
date: "6/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# README:  

This program is made to assist in pulling data from the team websites. It should be straightforward to adjust for different teams. 

For Spring Grove: 
-Input the list of URL's (manually enter them, or use an excel spreadsheet or something with all of them saved)
-Make sure the list of players who use HeadSet is up to date

For other teams:
-Check the above (list of URL's and Players)
-Adjust the html table reading functions. The info will likely be stored in differently numbered tables, etc.
-Adjust the column names when retrieving data from the website
-Adjust any data manipulation conducted after data retrieval


```{r Libraries}
library(tidyverse)
library(rvest)
library(stringr)
library(here)
```

```{r Lists}
# List of URLs to extract data from
URLList <- read.table(here("URLList.txt"))
PlayerList <- read.table(here("PlayerList.txt")) %>% 
  unite(Name, V1:V2, sep = " ", remove = TRUE, na.rm = TRUE)
```

```{r Helper_Functions}
# Determine if Spruce Grove is the home team or not
HomeTeam <- function(tables){
  Team <- tables %>%
  # Choose, from all the tables, which one we want
  # The . puts the pipe operating data set in front of the table number
  .[2] %>%
  # Create the R dataset from the webpage contents of the specified table
  html_table() %>%
  # This is just a thing we need to include, don't ask why
  .[[1]]

  # Determine if Spruce Grove is Home or Away  
  # If the chosen cell contains the Spruce Grove Saints, then they are the away team
  if (Team[2, 1] == "Spruce Grove Saints"){
    return(FALSE)
  }
  else {
    return(TRUE)
  }
}
# Return player stats tibble when Spruce Grove is home
HomePlayerStats <- function(tables){
  HomePlayerStats <- tables %>%
  # Home team player stats are in table 10 
  .[length(tables) - 2] %>%
  # Create the R dataset from the webpage contents of the specified table
  html_table() %>%
  # This is just a thing we need to include, don't ask why
  .[[1]] %>%
  # Remove the first two rows
  slice(c(-1, -2)) %>%
  # Rename the columns
  rename_all(~c("#", "Player", "G", "A", "PTS", "+/-", "SH", "PIM")) %>%
  # Make the dataset a tibble
  as_tibble()
  # Return the table of player stats
  return(HomePlayerStats)
}
# Return goalie stats tibble when Spruce Grove is home
HomeGoalieStats <- function(tables){
  HomeGoalieStats <- tables %>%
  # Choose, from all the tables, which one we want
  # Home team goalie stats are in table 12
  .[length(tables) - 1] %>%
  # Create the R dataset from the webpage contents of the specified table
  html_table() %>%
  # This is just a thing we need to include, don't ask why
  .[[1]] %>%
  # Remove the first two rows
  slice(c(-1, -2)) %>%
  # Rename the columns
  rename_all(~c("#", "Player", "Min", "SH", "SV")) %>%
  # Make the dataset a tibble
  as_tibble()
  
  # Remove any extra info after the goalie's name
  HomeGoalieStats$Player <- sub("^(\\S*\\s+\\S+).*", "\\1", HomeGoalieStats$Player)
  # Return the table fo the goalie stats
  return(HomeGoalieStats)
}
# Return player stats tibble when Spruce Grove is away
AwayPlayerStats <- function(tables){
  AwayPlayerStats <- tables %>%
  # Away team player stats are in table 11 
  .[length(tables) - 4] %>%
  # Create the R dataset from the webpage contents of the specified table
  html_table() %>%
  # This is just a thing we need to include, don't ask why
  .[[1]] %>%
  # Remove the first two rows
  slice(c(-1, -2)) %>%
  # Rename the columns
  rename_all(~c("#", "Player", "G", "A", "PTS", "+/-", "SH", "PIM")) %>%
  # Make the dataset a tibble
  as_tibble()
  # Return the table of player stats
  return(AwayPlayerStats)
}
# Return goalie stats tibble when Spruce Grove is away
AwayGoalieStats <- function(tables){
  AwayGoalieStats <- tables %>%
  # Choose, from all the tables, which one we want
  # Home team goalie stats are in table 12
  .[length(tables) - 3] %>%
  # Create the R dataset from the webpage contents of the specified table
  html_table() %>%
  # This is just a thing we need to include, don't ask why
  .[[1]] %>%
  # Remove the first two rows
  slice(c(-1, -2)) %>%
  # Rename the columns
  rename_all(~c("#", "Player", "Min", "SH", "SV")) %>%
  # Make the dataset a tibble
  as_tibble()
  
  # Remove any extra info after the goalie's name
  AwayGoalieStats$Player <- sub("^(\\S*\\s+\\S+).*", "\\1", AwayGoalieStats$Player)
  # Return the table fo the goalie stats
  return(AwayGoalieStats)
}
# Return the date of the match played
MatchDate <- function(tables){
  MatchDateTable <- tables %>%
  # The date is in table 7 
  .[7] %>%
  # Create the R dataset from the webpage contents of the specified table
  html_table() %>%
  # This is just a thing we need to include, don't ask why
  .[[1]]
  # Format the date for easier use in rearranging datasets
  MatchDate <- str_remove(MatchDateTable[2,2], "[.]") %>%
    str_remove(",") %>% 
    as.Date(MatchDate, format = "%b %d %Y")
  return(MatchDate)
}
```

```{r Extract_Player_And_Goalie_Stats}
# Extract the player and goalie stats for one game and return them in a list
PlayerGoalieStats <- function(URL){
  # Extract the HTML Tables
  HTMLtables <- read_html(URL) %>%
    html_nodes("table")
  # Determine if Spruce Grove is home/away
  Home <- HomeTeam(HTMLtables)
  # Grab data from correct table
  # If Spruce Grove is the HOME team
  if(Home == TRUE){
    PlayerStats <- HomePlayerStats(HTMLtables)
    GoalieStats <- HomeGoalieStats(HTMLtables)
  }
  # If Spruce Grove is the AWAY team
  else {
    PlayerStats <- AwayPlayerStats(HTMLtables)
    GoalieStats <- AwayGoalieStats(HTMLtables)
  }
  MatchDate <- MatchDate(HTMLtables)
  PlayerStats <- PlayerStats %>%
    add_column(Date = MatchDate, .after = "Player")
  GoalieStats <- GoalieStats %>%
    add_column(Date = MatchDate, .after = "Player")
  TeamStats <- list(PlayerStats, GoalieStats)
  return(TeamStats)
}
```

```{r Team_Stats}
# For each URL, grab the data tables for the players and the goalies)
for (URL in URLList$V1) {
  # Get the Player and Goalie Stats
  TeamStats <- PlayerGoalieStats(URL)
  # If it's the first URL, make the data set with that match's data
  if (URL == URLList$V1[1]){
    FullPlayerStats <- TeamStats[[1]]
    FullGoalieStats <- TeamStats[[2]]
    next
  }
  # For all other matches, combine the new data with the full dataset and keep only the data of players who use HeadSet
  FullPlayerStats <- bind_rows(FullPlayerStats, TeamStats[[1]])
  FullPlayerStats <- FullPlayerStats[which(FullPlayerStats$Player %in% PlayerList$Name), ]
  FullGoalieStats <- bind_rows(FullGoalieStats, TeamStats[[2]])
  FullGoalieStats <- FullGoalieStats[which(FullGoalieStats$Player %in% PlayerList$Name), ]
}

# Reorder the tables by player then by date
FullPlayerStats <- FullPlayerStats %>% 
  arrange(Player, Date)
FullGoalieStats <- FullGoalieStats %>% 
  arrange(Player, Date)

# Export the data tables
write.csv(FullPlayerStats, file = "SpruceGrovePlayerStats.csv")
write.csv(FullGoalieStats, file = "SpruceGroveGoalieStats.csv")
```
